---
title: "Campaign Investment Analysis"
author: Somya Verma
format: 
  html:
    theme:
      light: minty
      dark: cyborg 
  pdf: default
execute:
  echo: false
  warning: false
---
```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
library(lubridate)
library(stringr)
library(kableExtra)
library(dplyr)
```
# Introduction
#### Online retailers often invest in personalised marketing strategies to increase customer engagement and sales. However, it is not always clear whether higher marketing spend per customer leads to greater purchase amounts. This study aims to analyse customer-level data, including marketing spend, visits, and purchase behaviour, to understand this relationship. The findings can help businesses make more informed decisions about how to allocate their marketing budget effectively.

```{r}
#| echo: FALSE
AdCampaign_Data <- read.csv("36705535_Somya Verma.csv")
AdCampaign_Data <- AdCampaign_Data |> rename(
    Customer_ID = customer_id,
    Signup_Date= signup_date,
    Marketing_Expense= marketing_spend ,
    Num_Visits= visits,          
    Purchase_Amount= purchase_amount ,
    Customer_Type = segment,        
    Comments = notes 

)
View (AdCampaign_Data)
```
```{r}
#| echo: FALSE
#| results: hide
AdCampaign_Data <- AdCampaign_Data |> mutate(
    Comments=ifelse(Comments %in% c("","N/A"), NA, str_to_sentence(Comments))) |> 
    mutate(
    Purchase_Amount= ifelse(is.na(Purchase_Amount), 0.00, Purchase_Amount)) |>
    mutate(
       Signup_Date = parse_date_time(
      Signup_Date,
      orders = c("ymd", "dmy", "mdy", "ymd HMS", "dmy HMS", "b d, Y")
    ) |> as_date()
  )
glimpse(AdCampaign_Data)
```
# What are we looking at? (Research Question)
#### We will see if the two, dependent variable "Purchase" (by the consumers) and independent variable "Marketing Expense" (by the company on marketing) are correlated.

#### It will also include looking at the average purchase v/s the average marketing expense, and which segment of customers purchased the maximum. 

# Data Set Introduction

#### The dataset contains customer-level information from a personalised marketing campaign conducted by an online retailer. This dataset consists of `r nrow(AdCampaign_Data)` observations and `r ncol(AdCampaign_Data)` variables. Such as: 

* #### Customer ID
* #### Signup dates
* #### Marketing expense
* #### Number of visits
* #### Purchase Amount
* #### Customer type
* #### Comments


#### This data is used to analyse how marketing efforts influence customer purchasing behaviour. To prepare the dataset for analysis, several cleaning steps were performed. 

#### Column names were standardised for clarity, missing or inconsistent values in the comments field were handled, missing purchase amounts were replaced with 0, and date values were converted into a consistent format using the lubridate library. These steps ensured the dataset is clean, structured, and suitable for analysis.

# Satistical Analysis of Data

```{r}
#| echo: FALSE
#| label: tbl-summary
#| tbl-cap: "Summary Statistics of Key Variables"

Top_customer <- AdCampaign_Data |> 
  group_by(Customer_Type) |> 
  summarise(avg_purchase = mean(Purchase_Amount, na.rm = TRUE)) |> 
  arrange(desc(avg_purchase)) |> 
  slice(1)

summary_table <- AdCampaign_Data |> 
  summarise(
    Avg_marketing = mean(Marketing_Expense, na.rm = TRUE),
    Avg_purchase = mean(Purchase_Amount, na.rm = TRUE),
    Avg_visits = mean(Num_Visits, na.rm = TRUE)
  )
  
kable(summary_table, caption = "Summary Statistics of Key Variables") |>
  kable_styling(full_width = FALSE, position = "center") |> 
  column_spec(1:3, width = "4cm")

```
##### The summary statistics in @tbl-summary provide an overview of customer behaviour in the dataset. On average, the _**marketing expense per customer is 154.38**_, while the _**average purchase amount is 131.78**_, indicating that customers generate substantial purchases in response to marketing efforts. Customers visit the platform approx 5.76 times on average, suggesting a good level of engagement. <br>

```{r}
#| echo: FALSE
#| label: tbl-Customer_summary
#| tbl-cap: "Summary of Customer Segment"
customer_summary <- AdCampaign_Data |>
  group_by(Customer_Type) |>
  summarise(
    Total_Customers = n(),
    Avg_Purchase = mean(Purchase_Amount, na.rm = TRUE)
  ) |>
  arrange(desc(Avg_Purchase))

kable(customer_summary, caption = "Customer Distribution and Average Purchase by Type")

```

##### Among the different customer segments, the _**Consumer group**_ has the highest average purchase amount, indicating that this segment contributes most to overall sales. These insights provide useful context for understanding customer behaviour before analysing the relationship between marketing expense and purchase amount.

```{r}
#| echo: FALSE 
#| results: hide
model <- lm(Purchase_Amount ~ Marketing_Expense, data = AdCampaign_Data)

summary(model)
```
## Relationship between Purchase & Expense.

##### The @fig-plot1 plot shows a clear positive relationship between marketing expense and purchase amount, with higher spending associated with increased customer purchases. The linear regression model confirms this, explaining approximately 71.9% of the variation in purchase amount (R² = 0.719).

##### The relationship is highly statistically significant (p < 0.001), indicating a reliable effect of marketing on sales. Despite some variability (residual standard error ≈ 49.75), the results suggest that marketing expense is a strong predictor of customer purchasing behaviour.
```{r}
#| echo: FALSE
#| result: hide
#| label: fig-plot1
#| fig-cap: "Relationship: marketing expense and purchase amount"
sorted_data <- AdCampaign_Data %>%
  arrange(Marketing_Expense)

ggplot(sorted_data, aes(x = Marketing_Expense, y = Purchase_Amount)) +
  geom_point(alpha = 0.6)+
  geom_line(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,
              linewidth = .5)+
  labs(
    title = "Marketing Expense vs Purchase Amount",
    x = "Marketing Expense",
    y = "Purchase Amount",
    caption = "Relationship: marketing expense and purchase amount", 
  )+theme(
    plot.title = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5)
  )

```
# Final Result: Do We Still Spend On Marketing? 
#### The answer is a Yes. There are few things that we must give consideration to: <br>

* #### **Satistical Outcome**: Considering @tbl-summary, one might say that the average expense is more than purchase, however, the difference however is very less, and business growth is a long process, hence the gap will easily be covered to move from breakeven to profit points for the company. Average visits also indicate that customers are engaging with the advertisements. The @fig-plot1 suggests a strong a relationship and impact of marketing expense on the purchase. <br>

* #### **Customer Segments**: As seen in the @tbl-Customer_summary, there are various customer segments. The company should rather bifurcate it's strategies of marketing based on the segment of customers the campaign or advertisement is made to target. <br>

* #### **Other Factors**: Business growth depends on many other factors than just Marketing, and hence, while it does play a major role, brand loyalty, age of the business, seasonality, and target audience also have effects on why the customer would purchase. <br>

# Limitations of Data

* #### **Missing values**: There were certain missing values and string data such as Comments are not exactly recorded consistently for the data set. Date records too were found to be missing.  

* #### **Mixed pattern**: While date section had various formats that was cleaned, empty cells under purchase amount or "NA" was recorded. 
